%%%%%%%%%%%%%
% 
% Alexander Powell
% Finite Automata
% Homework Assignment #5
% 10.15.2015
% 
%%%%%%%%%%%%%

\documentclass[11pt]{article}

\usepackage{times,mathptm}
\usepackage{pifont}
\usepackage{exscale}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{tikz}


\textwidth 6.5in
\textheight 9in
\oddsidemargin -0.0in
\topmargin -0.0in

\parindent 0pt     % How much the first word of a paragraph is indented. 
\parskip 0pt	   % How much extra space to leave between paragraphs.

\begin{document}

\begin{center}             % If you only centering 1 line use \centerline{}
\begin{LARGE}
{\bf Finite Automata Homework 5}
\end{LARGE}
\vskip 0.25cm      % vertical skip (0.25 cm)

Due: Thursday, Oct 15\\  % force new line
Alexander Powell
\end{center}

\begin{enumerate}

\item % Problem 1
(a) Let $B = \{ 1^ky | y \in \{ 0,1 \}^* \}$ and $y$ contains at least $k$ $1s$ for $k \geq 1$.  Prove $B$ is regular.  

\textbf{Solution: }
\begin{proof}

To prove $B$ is regular we simply need to create a regular expression that generates all words in $B$.  We know that all words need to start with a $1$ and they need to have at least one more $1$ somewhere in them.  The regex is shown below:
$$ (1)(0)^*(1)(0 \cup 1)^* $$

\end{proof}

(b) Let $C = \{ 1^ky | y \in \{ 0,1 \}^* \}$ and $y$ contains at most $k$ $1s$ for $k \geq 1$.  Prove $C$ is not regular.  

\begin{proof}

To prove $C$ is not regular we can use the pumping lemma for regular languages.  Begin by assuming that $C$ is indeed a regular language, then the pumping lemma applies.  We can choose our $s$ to be $s = 1^P01^p$, $s = xy^iz$, $|xy| \leq P$, and $|y| \neq 0$ where $P$ is our pumping length.  From this, we can see the string $xy$ in $s$ is composed only of $1s$ since its length is less than $P$.  Therefore, we can write $$s = 1^k1^m1^{P-k-m}01^P,$$ where $x = 1^k$, $y = 1^m$, and $z = 1^{P-k-m}01^P$.  

By taking $i = 0$ so $xy^0z = 1^k1^{P-k-m}01^P = 1^{P-k}01^P$.  Here we have the number of $1s$ in the second part of the string is larger than the number of $1s$ in the first part, so $xy^0z$ is not an element of $C$.  This is a contradiction to our assumption, and therefore $C$ is not a regular language.  

\end{proof}

\item % Problem 2
Give the context free grammar for the complement of the language
$$ A = \{ a^nb^n | n \geq 0 \} $$

\textbf{Solution: }
The given language contains all words with some number of $a$s followed by an equal number of $b$s.  Therefore, the complement can be described as the language of all strings where the number of $a$s and $b$s differ as well as all strings not in the pattern of $a^ib^j$.  This can be written as:
$$ \bar{A} = \{ a^ib^j | i > j \} \cup \{ a^ib^j | i < j \} \cup (a \cup b)^*b(a \cup b)^*a(a \cup b)^* $$
Then the union of these languages can be represented with the grammar:
$$ S \rightarrow S_1 | S_2 | S_3 $$
The language $\{ a^ib^j | i > j \}$ can be represented with:
$$ S_1 \rightarrow aS_1b | aS_1 | a $$
and the language $\{ a^ib^j | i < j \}$ can be represented with:
$$ S_2 \rightarrow aS_2b | S_2b | b $$
The regular expression is represented with 
$$ S_3 \rightarrow S_4bS_4aS_4 $$
$$ S_4 \rightarrow aS_4 | bS_4 | \epsilon $$

Finally, we can write the context free grammar as a whole with:
$$ S \rightarrow S_1 | S_2 | S_3 $$
$$ S_1 \rightarrow aS_1b | aS_1 | a $$
$$ S_2 \rightarrow aS_2b | S_2b | b $$
$$ S_3 \rightarrow S_4bS_4aS_4 $$
$$ S_4 \rightarrow aS_4 | bS_4 | \epsilon $$


\item % Problem 3

\textbf{Solution:}
\begin{enumerate}

\item
The language of this grammar can be defined as the union of two sets.  All strings generated from $T$ can be represented as the language
$$ S_1 = \{ 0^i\#0^j | i,j \geq 0 \}, $$
and all strings generated from $U$ can be represented as the language
$$ S_2 = \{ 0^i\#0^{2i} | i \geq 0 \} .$$
The language can be written as $S_1S_1 \cup S_2$.  

\item
To prove $L(G)$ is not regular we will use closure properties.  Begin by assuming that $L(G)$ is a regular language.  Take the homomorphism $H_1$ where $H_1(0) = 0$, $H_1(\#) = \#$, and $H_1(1) = 00$.  Also let $M_1 = H_1^{-1}(L(G))$ and $M_2 = M_1 \cap (0^*\#1^*)$.  Then we can write $M_2 = \{ 0^i\#1^i | i \geq 0 \}$.  

Now, let's take another homomorphism $H_2$ where $H_2(0) = 0$, $H_2(1) = 1$, and $H_2(\#) = \epsilon$ and $M_3 = H_2(M_2) = \{ 0^n1^n | n \geq 0 \}$.  

Since $M_1, M_2$, and $M_3$ are obtained through $L(G)$, using closure properties that preserve regularity, and we see that $M_3$ is clearly a non-regular language, we have a contradiction to our original assumption.  Therefore $L(G)$ is not regular.  

\end{enumerate}

\item % Problem 4
Give the context free grammar for the language
$$ L = \{ a^ib^jc^k | i + j \neq k \} $$

\textbf{Solution: }
It's easiest to begin by creating a grammar that generates the complement of the language (ie. when $i+j = k$).  To start, we can write the two production rules 
$$ C \rightarrow aCc | B $$
$$ B \rightarrow bBc | \epsilon $$
\newpage
Now, to change this grammar to generate words in the language where $i+j \neq k$ we can add the following production rules to ensure cases where $i+j < k$
$$ S_1 \rightarrow Lc $$
$$ L \rightarrow Lc | C $$
Now, we need to create production rules when $i+j > k$, shown below.  
$$ S_2 \rightarrow R_bb | R_aa $$
$$ R_a \rightarrow R_aa | R_b | C $$
$$ R_b \rightarrow R_bb | B $$

By putting all the rules together to form one context free grammar, we get:
$$ S \rightarrow S_1 | S_2 $$
$$ S_1 \rightarrow Lc $$
$$ L \rightarrow Lc | C $$
$$ S_2 \rightarrow R_bb | R_aa $$
$$ R_a \rightarrow R_aa | R_b | C $$
$$ R_b \rightarrow R_bb | B $$
$$ B \rightarrow bBc | \epsilon $$
$$ C \rightarrow aCc | B $$


\end{enumerate}

\end{document}









