%%%%%%%%%%%%%
% 
% Alexander Powell
% Finite Automata
% Homework Assignment #10
% 11.19.2015
% 
%%%%%%%%%%%%%

\documentclass[11pt]{article}

\usepackage{times,mathptm}
\usepackage{pifont}
\usepackage{exscale}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{array}
\usepackage{lipsum}


\textwidth 6.5in
\textheight 9in
\oddsidemargin -0.0in
\topmargin -0.0in

\parindent 0pt     % How much the first word of a paragraph is indented. 
\parskip 0pt	   % How much extra space to leave between paragraphs.

\begin{document}

\begin{center}             % If you're only centering 1 line use \centerline{}
\begin{LARGE}
{\bf Finite Automata Homework 10}
\end{LARGE}
\vskip 0.25cm      % vertical skip (0.25 cm)

Due: Thursday, Nov 19\\  % force new line
Alexander Powell
\end{center}

\begin{enumerate}

\item Prove that $A = \{ w_{2i}\text{ }|\text{ }w_{2i} \not \in L(M_i) \}$ is not Turing-recognizable, where $w_{2i}$ is the $2i$th string in the lexicographicorder of binary strings and $M_i$ is the TM whose binary code is $w_i$.

\textbf{Solution: }
\begin{proof}

If $A = \{ w_{2i}\text{ }|\text{ }w_{2i} \not \in L(M_i) \}$, then we can also write $A = \{ <M>|M \text{ is a TM which doesn't accept }<M> \}$.  Let's assume that $A$ is Turing recognizable.  Then there exists a TM $M$ such that $A = L(M_i)$, for each $i$.  From the definition of $A$, we know $w_{2i} \in A$ when $w_{2i} \not \in L(M_i)$, which is the case when $w_{2i} \in L(M_i)$.  This is a contradiction to our assumption.  Therefore, we can conclude that $A$ is not Turing recognizable.  

\end{proof}

\item Prove that $ES_{TM} = \{ \text{code of M: TM M accepts the empty string} \}$ is undecidable. (Hint: Reduction from $A_{TM}$.)

\textbf{Solution: }
\begin{proof}

We begin by reducing from $A_{TM}$.  Let's assume, to the contrary, that $M$ can decide $ES_{TM}$.  Then we can construct a Turing machine $D$ that uses $M$ to show $A_{TM}$ is Turing decideable.  The machine $D$ works as follows:
$D$ takes in the input $<M,w>$.  Inside $D$, the machine will process each string.  $M$ runs on $w$; if $M$ accepts $w$, then $D$ accepts, else $D$ rejects.  

Since this machine decides $A_{TM}$, we have a contradiction to the statement that $A_{TM}$ is non Turing decideable.  
Therefore, we can conclude that our assumption is false and this $ES_{TM}$ is not decideable.  

\end{proof}

\item Let $T = \{<M>\text{ }|\text{ }M\text{ is a TM that accepts }w^R\text{ whenever it accepts w }\}$. Show that $T$
is undecidable.  

\textbf{Solution: }
\begin{proof}

We will use a proof by contradiction.  Let's assume, to the contrary, that $T$ is decideable.  Then there exists some Turing machine $D$ that decides $T$.  We can reduce $A_{TM}$ to $T$, where 
$$ A_{TM} = \{ <M,w>\text{ }|\text{ }M \text{ is a TM that accepts }w \}, $$
and we know $A_{TM}$ is undecidable.  Also, we will denote the machine that accepts $A_{TM}$ as $A$.  The description of $A$ is as follows:\\
It will take in an input of $<M,w>$ ($M$ is a Turing machine and $w$ is a string).  
\begin{enumerate}
\item First, $A$ will contruct TM $R$.  On input $x$: 
\item If $x$ is in a reversible form, then $R$ accepts.  If not, then $w$ is run through $R$ and either accepts of rejects.  
\item Then $T$ is run on the code of $<R>$.  If $T$ accepts then $A$ accepts, and if $T$ rejects then $A$ rejects.  
\end{enumerate}
Thus, we have shown that $A_{TM}$ is decidable, which is a contradiction to our original assumption.  Therefore, we can conclude that $T$ is undecidable.  
\end{proof}


\end{enumerate}

\end{document}




































