%%%%%%%%%%%%%
% 
% Alexander Powell
% Finite Automata
% Homework Assignment #7
% 10.29.2015
% 
%%%%%%%%%%%%%

\documentclass[11pt]{article}

\usepackage{times,mathptm}
\usepackage{pifont}
\usepackage{exscale}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{tikz}


\textwidth 6.5in
\textheight 9in
\oddsidemargin -0.0in
\topmargin -0.0in

\parindent 0pt     % How much the first word of a paragraph is indented. 
\parskip 0pt	   % How much extra space to leave between paragraphs.

\begin{document}

\begin{center}             % If you're only centering 1 line use \centerline{}
\begin{LARGE}
{\bf Finite Automata Homework 7}
\end{LARGE}
\vskip 0.25cm      % vertical skip (0.25 cm)

Due: Thursday, Oct 29\\  % force new line
Alexander Powell
\end{center}

\begin{enumerate}

\item
\begin{enumerate}
\item

Use the pumping lemma to show that $A = \{ 0^n1^n0^n1^n | n \geq 0 \}$ is not a context free language.  

\textbf{Solution: }

\begin{proof}

Assume $A$ is a context free language.  Then the pumping lemma for context free languages applies to $A$.  If we take $s = 0^p1^p0^p1^p$ then clearly $|s| = 4p \geq p$.  Also, we can describe say $s = uvxyz$ where $|vxy| \leq p$.  In other words, $s$ can be written as:
$$s = \underbrace{0 \ldots 0}_\text{p}\underbrace{1 \ldots 1}_\text{p}\underbrace{0 \ldots 0}_\text{p}\underbrace{1 \ldots 1}_\text{p}$$
This creates three cases:
\begin{enumerate}
\item\textbf{Case 1: } $vxy$ is composed of all $0s$ and is therefore contained completely within the first or second string of $0s$ of length $p$.  Because we know that $|vy| > 0$ then either $v$ or $y$ must have at least one $0$.  By setting $i = 0$ we get $s = uv^0xy^0z$, which causes one of the two strings of $0s$ in the original $s$ to have at least one fewer $0s$ than the other.  This is a contradiction to our assumption because $uv^0xy^0z \not \in A$.  

\item\textbf{Case 2: } $vxy$ is composed of all $1s$ and thus is made up of either the first or second string of $1s$ in the original $s$.  Now, either $v$ or $y$ must contain at least one $1$ since $|vy| > 0$.  Again, we set $i=0$ and get $uv^0xy^0z$ which generates a string where one of the two substrings of $1s$ has a different length than the other (by at least one).  Therefore $uv^0xy^0z \not \in A$, which is a contradiction to our assumption.  

\item\textbf{Case 3: } The third case covers scenarios where $vxy$ is made up of a combination of $0s$ and $1s$.  We can further split this into two cases:
\begin{enumerate}
\item\textbf{Subcase 3a: } $vxy$ is made up of $0s$ followed by $1s$

In this case, either $vxy$ straddles the first partition of $0s$ and $1's$ or the second partition.  Because $|vxy| \leq p$ then by ``pumping" up or down the lengths of the substrings of $0s$ and $1s$ immediately next to $vxy$ will be changed but not the other substring.  If we fix $i = 2$, then $uv^2xy^2z$ would give us a string like $0^{p'}1^{p'}0^p1^p$ or $0^p1^p0^{p'}1^{p'}$, where $p' \neq p$.  This is a contradiction to our assumption because these strings are not elements of $A$.  

\item\textbf{Subcase 3b: } $vxy$ is made up of $1s$ followed by $0s$

The same argument can be made for the case where a string is made up of $1s$ followed by $0s$.  Again, let's fix $i = 2$, then $uv^2xy^2z$ gives us a string that looks like $0^p1^{p'}0^{p'}1^p$ which is clearly not in $A$ because $p' \neq p$.  Again, we have a contradiction.  
\end{enumerate}

So, because we have shown that we reach a contradiction in every possible permutation of $s$, we have proven that $A$ is not a context free language.  
\end{enumerate}
\end{proof}

\newpage
\item

Use the pumping lemma to prove that 
$$D = \{ t_1\#t_2\#\ldots\#t_k|k \geq 2, \text{each } t_i \in \{ a,b \}^* \text{ and } t_i = t_j \text{ for some } i \neq j \}$$
is not a context free language.  

\textbf{Solution: }
\begin{proof}

Assume to the contrary that $D$ is a context free language so the pumping lemma applies.  We can take $s = a^pb^p\#a^pb^p \in D$.  Then we have that $|s| = 4p > p$.  Also, we know that $|vxy| \leq p$ and $|vy| > 0$.  

\begin{enumerate}
\item\textbf{Case 1: } If $vxy$ is contained entirely on either the left or right side of the $\#$.  

By setting $i = 2$ then $uv^2xy^2z$ will cause $s$ to look like: 
\[
\begin{gathered}
a^{p'}b^p\#a^pb^p \text{ or, } \\
a^pb^{p'}\#a^pb^p \text{ or, } \\
a^pb^p\#a^{p'}b^p \text{ or, } \\
a^pb^p\#a^pb^{p'} \text{     }
\end{gathered}
\]
where $p' \neq p$.  Clearly none of these strings is in the language $D$ so we have a contradiction.  

\item\textbf{Case 2: } If $vxy$ contains the $\#$ character.  
\begin{enumerate}
\item\textbf{Subcase 2a: } If $v$ or $y$ contains the $\#$ then we can set $i=0$ to get $uv^0xy^0z = uxz$ which doesn't contain the $\#$ character so $uv^0xy^0z$ is not an element of the language $D$, which is a contradiction.  

\item\textbf{Subcase 2b: } If $x$ contains the $\#$ character then $v$ is a substring of $b^p = underbrace{b \ldots b}_\text{p}$ and $y$ is a substring of $a^p = underbrace{a \ldots a}_\text{p}$.  Again, by setting $i=0$ then we get $uv^0xy^0z$ which looks something like $a^pb^{p'}\#a^{p'}b^p$ where $p' \neq p$.  Therefore, $t_1 \neq t_2$ so $uv^0xy^0z \not \in D$, which is a contradiction to our original assumption.  
\end{enumerate}
\end{enumerate}
Since we've ended up with contradictions at every possible case of representing $s$, we can conclude that $D$ is not a context free language.  
\end{proof}

\end{enumerate}

\newpage
\item

Let $\Sigma = \{ 1,2,3,4 \}$ and $C = \{ w \in \Sigma^* | \text{ in w, the number of }1s\text{ equals the number of }2s\text{ and the number of }3s\text{ equals the number of }4s. \}$  Show that $C$ is not a context free language.  

\textbf{Solution: }
\begin{proof}
Assume to the contrary that $C$ is a context free language.  Then the pumping lemma for context free languages applies to $C$.  We can divide this proof up into the four cases shown below.  
\begin{enumerate}
\item\text{Case 1: } $vxy$ contains a $1$

By setting $i=2$, then $uv^2xy^2z \not \in C$ because $|vxy| \leq p$ so the number of $1s$ and $2s$ will be different.  This generates a string that's not in the language $C$, so we have a contradiction.  

\item\text{Case 2: } $vxy$ contains a $2$

The rest of the cases are similar to the first.  By setting $i=2$, then $uv^2xy^2z \not \in C$ because the number of $1s$ and $2s$ will be different.  This generates a string that's not in the language $C$, so again we have a contradiction.  

\item\text{Case 3: } $vxy$ contains a $3$

Again, by setting $i=2$, then $uv^2xy^2z \not \in C$ because the number of $3s$ and $4s$ will be different.  This generates a string that's not in the language $C$, so again we have a contradiction.  

\item\text{Case 4: } $vxy$ contains a $4$

Again, by setting $i=2$, then $uv^2xy^2z \not \in C$ because the number of $3s$ and $4s$ will be different.  This generates a string that's not in the language $C$, so again we have a contradiction.  
\end{enumerate}
Because we end up with a contradiction in every step we have proven that $C$ is not a context free language.  
\end{proof}

\end{enumerate}
\end{document}




































